---
title: "Statistical_Inference_1"
author: "Ajith Masthan"
date: "May 17, 2020"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

# Statistical Inference Part-1

## Overview
Explore the exponential distribution in R and compare it with the Central Limit Theorem. 
The exponential distribution is simulated in R with rexp(n, lambda) where lambda is the rate parameter. 
The mean and standard deviation of exponential distribution is 1/lambda. 
lambda = 0.2 is used for all of the simulations. 
Distribution of averages of 40 exponentials is investigated with a thousand simulations. 


## Simulations

```{r}
# load neccesary libraries
library(ggplot2)

# set constants
lambda <- 0.2 # lambda for rexp
n <- 40 # number of exponetials
numberOfSimulations <- 1000 # number of tests

# set the seed to create reproducability
set.seed(2866)

# run the test resulting in n x numberOfSimulations matrix
exponentialDistributions <- matrix(data=rexp(n * numberOfSimulations, lambda), nrow=numberOfSimulations)
exponentialDistributionMeans <- data.frame(means=apply(exponentialDistributions, 1, mean))
```

```{r echo=FALSE}
# plot the means
ggplot(data = exponentialDistributionMeans, aes(x = means)) + 
  geom_histogram(binwidth=0.1) +   
  scale_x_continuous(breaks=round(seq(min(exponentialDistributionMeans$means), max(exponentialDistributionMeans$means), by=1)))
```

## Sample Mean versus Theoretical Mean

The expected mean mu of a exponential distribution of rate lambda is 

mu = 1/lambda

```{r}
mu <- 1/lambda
mu
```

X_bar is the average sample mean of 1000 simulations of 40 randomly sampled exponential distributions.

```{r}
meanOfMeans <- mean(exponentialDistributionMeans$means)
meanOfMeans
```

The expected mean is close to average sample mean


## Sample Variance versus Theoretical Variance

The expected standard deviation sigma of a exponential distribution of rate lambda is 

sigma = (1/lambda)/sqrt(n)



```{r}
sd <- 1/lambda/sqrt(n)
sd
```

The variance Var of standard deviation sigma is

Var=square(sigma)

```{r}
Var <- sd^2
Var
```

Let Var_x is the variance of the average sample mean of 1000 simulations of 40 randomly sampled exponential distribution with standard deviation sigma_x
```{r}
sd_x <- sd(exponentialDistributionMeans$means)
sd_x

Var_x <- var(exponentialDistributionMeans$means)
Var_x
```

The Standard deviation and Variance are close

## Distribution

Population Mean & Standard Deviation with a normal distribution of the expected values. 

```{r echo=FALSE}
# Plotting the means
ggplot(data = exponentialDistributionMeans, aes(x = means)) + 
  geom_histogram(binwidth=0.1, aes(y=..density..), alpha=0.2) + 
  stat_function(fun = dnorm, arg = list(mean = mu , sd = sd), colour = "red", size=1) + 
  geom_vline(xintercept = mu, size=1, colour="#CC0000") + 
  geom_density(colour="blue", size=1) +
  geom_vline(xintercept = meanOfMeans, size=1, colour="#0000CC") + 
  scale_x_continuous(breaks=seq(mu-3,mu+3,1), limits=c(mu-3,mu+3)) 
```

Observarion : The calculated distribution of means of random sampled exponantial distributions matches with the normal distribution with the expected values based on the given lamba
