---
title: "Statistical Inference 2"
author: "Ajith Masthan"
date: "May 17, 2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Statistical Inference Part-2

## Overview
ToothGrowth data is loaded and some basic exploratory data analyses is peformed
Basic Summarixations of the data is done
Confidence intervals and/or hypothesis tests are used to compare tooth growth by supp and dose.
The finally few conclusions are made along with assumptions needed for those conclusions. 


## Data Import

```{r}
# Libraries
library(ggplot2)
library(datasets)
library(gridExtra)
library(GGally)

# Vitamin C effect on Tooth Growth in Guinea Pigs
data(ToothGrowth)
toothGrowth <- ToothGrowth 
toothGrowth$dose <- as.factor(toothGrowth$dose)
```

## Summaries
```{r}
str(toothGrowth)
summary(toothGrowth)
head(toothGrowth)

table(toothGrowth$supp, toothGrowth$dose)
```
```{r echo=FALSE}
p1 <- ggplot(data=toothGrowth, aes(x=dose,y=len,fill=dose)) +
  geom_boxplot() + 
  theme(legend.position="none") + 
  facet_grid(.~supp) 

p2 <- ggplot(data=toothGrowth, aes(x=supp,y=len,fill=supp)) +
  geom_boxplot() + 
  theme(legend.position="none") + 
  facet_grid(.~dose) 

p3 <- ggplot(data=toothGrowth, aes(x=supp,y=len,fill=supp)) +
  geom_boxplot()

p4 <- ggplot(data=toothGrowth, aes(x=dose,y=len,fill=dose)) +
  geom_boxplot()

grid.arrange(p1, p4, p2, p3, ncol = 2, nrow=2)

```

Analysis of Variance (ANOVA)
```{r}
anova.out <- aov(len ~ supp * dose, data=toothGrowth)
summary(anova.out)
```

Observation: There is a relationship between the 
length (len) and dosage (dose) (F(1,54)=15.572;p<0.01)
length(len) and supplement type (supp) (F(2,54)=92;p<0.01). 
Combination of supplement type (supp) and dosage (dose) compared to the length (len) (F(2,54)=4.107;p<0.05).

```{r}
TukeyHSD(anova.out)
```

Observation : 
There are significant differences between each of the groups in supp and dose
Only the interactions between VC:0.5-OJ:0.5; VC:1-OJ:0.5; OJ:2-OJ:1; VC:2-OJ:1 and VC:2-OJ:2  are not significant
```{r}
confint(anova.out)
print(model.tables(anova.out,"means"),digits=3)
```

## Conclusions

Supplement and the dosage have clear individual effects on the length of teeth in guinea pigs. 
More the means on average longer teeth. 
Supplement type has a clear influence but OJ has a greater avarage teethgrowth in combination with dosages 0.5 and 1 then for the VC supplement.
While teeth length for the VC supplement vs the OJ in combiantion with dosage 2 has no significant effect (almost same mean & same confidence interval)

These conclusions are based on the assumptions of the facts that

* The guinea pigs are repesentative for the population of guinea pigs, 
* The dosage and supplement were randomly assigned and 
* The distribution of the means is normal.